---
title: "Class 21"
author: " "
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* Hypothesis tests and p-values (lecture)
* Group Work

Note: for smaller sample sizes (<30), we can use _qt()_ to obtain critical values and confidence intervals. This is the equivalent of _qnorm_ for the t-distribution.

```{r}

# qt( level, degrees of freedom)
```


We can generate t-statistics manually. However, the _t.test_ command does this automatically:

Let's test whether the callback rate for black resumes is different from 0, at the 99% level.


## One sample t-test (for a single distribution):

```{r}
library(tidyverse)
resume <- read.csv("data/resume.csv") %>% mutate(female=ifelse(sex=="female",1,0))

t.test(,conf.level=)

```

We can also test whether it is different from another value (such as 2%), using the ,mu option:

```{r}

t.test(,mu=,conf.level=)
```

## Two sample t-test  (t-test for difference in means):

Let's test whether the difference in callback rates for black and white resumes is statistically significant, at the 95% level

```{r}

t.test( , )

```

We can also use t-tests to check for accurate randomization. Let's check that female was independently randomized between the white and black resumes, at the 95% level:

```{r}

t.test(, )
```

We can also use regression to obtain t-statistics. Note that regression always tests the null hypothesis that the true coefficient is 0.

```{r}

result <- lm(,data=resume)
summary(result)

```

## p-values

What is the chance of seeing our t-statistic? In other words, how much data is outside the boundaries of our t-statistic?

We need to use the _pt_ function, which takes 3 arguments:

- The absolute value of the t-statistic
- The degrees of freedom (n-1)
- lower.tail=FALSE

```{r}
tstatistic <- abs(-1.33)
n <- 15

# p value
pt(tstatistic, df=n-1, lower.tail=FALSE)*2
```

If we are using a 90% confidence interval, what is alpha?

Do we reject or fail to reject our null hypothesis at the 90% level?



# Group Work: Practice with Hypothesis Testing 

Choose a study below, and determine whether we can reject the null hypothesis that the treatment effect is 0. If you finish a study, work on another one.


##A

Re-evaluate the Boston immigration experiment. Recall that the outcome in this study is a change in attitudes relating to immigration (diff_imm), while treatment indicates whether someone was exposed to immigrants on their daily train commute.


```{r}
boston <- read.csv("data/boston.csv")
boston$diff_imm <- boston$immigration_after - boston$immigration_before
```

Obtain a 90% confidence interval for the treated and control groups. Calculate the t-statistic and p-value (at the 90% level) for the null hypothesis that the difference between the treated group and the control group is 0. Do you reject or fail to reject the null?

```{r}

```

Would your answer change if you used a 95% confidence level?

```{r}

```

Next, evaluate whether age, income, white, college, and usborn are balanced between the treated and control groups (i.e., the difference in means is 0), using two sample t-tests and a 90% confidence interval. What is the associated p value for each background characteristic?

```{r}

```

##B

Re-evaluate the 9/11 victims study. 

voter.id	            Unique identifiers of relatives and neighbors of the victims
treatment           	Linked to victims (1) vs control group (0)
victim.status	        Type of link: Families (2) vs neighbors (3) 
ge2000	              Voting in the 2000 general election (Y=yes,at the polls, A=yes,absentee, N=No)
ge2002	              Voting in the 2002 general election (1=Yes, 0=No)
fam.members	          Number of family members living with voter at their address
age	                  Voter’s age
party	                Voter’s party affiliation (D=Democrat, R=Republicans, I=independent, N=no affiliation)
sex	                  Voter’s reported sex
pct.white	            Proportion of non-Hispanic white voters living on the same block
median.income	        Median income of voters living on the same block


```{r}
terrorism <- read.csv("data/terrorism.csv") 
```

Recall that the outcome in this study is voting in the 2002 elections (ge2002), while treatment indicates whether someone was personally linked to someone who died in the attacks. 

Calculate the average treatment effect, and obtain a 95% confidence interval for the ATE. Obtain a p-value, and then evaluate the null hypothesis that the true effect of a personal link on voting behavior is 0.

```{r}

```

Evaluate whether individuals in the treated and control group are comparable with respect to background characteristics (i.e., the difference in means is 0), using a 95% confidence level. The background characteristics are: ge2000, fam.members, age, party, sex, pct.white, median.income

```{r}

```


##C

Re-evaluate the Kenyan savings experiments.

```{r}
kenya <- read.csv("data/rosca2.csv")
kenya <- subset(kenya,has_followup2 == 1)
```

Variables:
fol2_amtinvest  Amount invested in Kenyan shillings, in Wave 2 (outcome)
encouragement   Binary variable indicating whether individuals are in the control group (1) or not (0)
safe_box        Binary variable indicating whether individuals are in the safe deposit box treatment group (1) or not (0)
locked_box       Binary variable indicating whether individuals are in the locked box treatment group (1) or not (0)
bg_b1_age       age of participant
bg_female       gender of participant
bg_married     marital status of participant


Calculate the t-statistic and p-value (at the 95% level) for the null hypothesis that the difference between the locked box group and the control group is 0. Repeat for the safe box group. What is the 95% confidence interval for the difference in means?
 
```{r}

```

Now evaluate balance between the treated and control conditions with respect to age, gender, and marital status, at the 95% confidence level. Note: since there are two treatment conditions, you will have to evaluate the balance of each treatment relative to control.

```{r}

```



## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?


