---
title: "Class 13"
author: " "
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* Review of Merging
* More Diagnostics
* Intro to Regression (Lecture)
* Regression in R

# Merging

How did the latest available poll in each state compare to the average of the last two weeks?

```{r}
library(tidyverse)
library(lubridate)

polls08 <- read.csv("data/polls08.csv")
polls08 <- polls08 %>% mutate(days.until=mdy("11/4/08") - mdy(middate))
polls08 <- polls08 %>% mutate(poll.margin = Obama - McCain)

pres08 <- read.csv("data/pres08.csv")
pres08 <- pres08 %>% mutate(act.margin = Obama - McCain)
```

Modify the code to average across all polls conducted _within the last two weeks_ before the election. Does the accuracy increase or decrease compared to looking at the most recent poll?

```{r}

```

## More Diagnostics 

Scatter Plot

```{r}
x<-
y<-
plot(x,y, xlab = "Poll Results", ylab = "Actual Election Results")
```

Add text:

```{r}
x<-
y<-
  
plot(type="n",x =  x , y = u, xlab = "Poll Results", ylab = "Actual Election Results")
text(x = x  , y = y, 
     labels = , 
     col = "blue",
     cex=.5)

# Diagnosis Lines
abline(0, 1, lty = "dashed")
abline(v=0)
abline(h=0)
```


# Regression (lecture)

## Regression in R

Load the Danish crime data:

```{r}
denmark <- read.csv("data/denmark.csv")
```

Fitting regressions in R is straightforward: we can use the function lm()

```{r}
y <- denmark$violentcrime_rate
x <- denmark$percap_spending_educ
lm()
```

To visualize, we can superimpose the regression line on the data using abline():

```{r}

```

Other regressions:

```{r}

```



# Remember to complete the survey (link posted in Slack)

## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?