---
title: "Class 2"
author: ""
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
library(tidyverse)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Class Plan

* Variables
* Vectors
* Intro to dataframes

# Variables

R can save information as a variable using the following syntax: variable name <- value.

Let's create a variable named output, with the value of 13.

```{r}

```

We can input variable names into functions. Let's take the natural log of *output*

Note: There are many helpful lists of functions online: https://www.statmethods.net/management/functions.html

```{r}

```

Get the value of the variable *output* by typing the variable name

```{r}

```

Let's change the value of *output* by overwriting it

```{r}

```

Now let's copy the variable *output* to a new variable called *output2*

```{r}

```

We can perform mathematical operations on variables

```{r}

```

## String (text) variables

We can also store text within variables by wrapping the text in double quotation marks (""). 
Let's create a new variable called output3 that contains the words "gov 10"

```{r}

```

We can't perform mathematical operations on text

```{r}

```

# Working with vectors 

*Vectors* are one-dimensional arrays.  You can think of them as a single column of values.  We can manually create a vector using the *c()* function.  With c(), we can concatenate objects into a single vector

Let's look at the schools with the largest undergraduate populations:

| School (state)                                | Fall 2015 undergraduate enrollment |
|-----------------------------------------------|------------------------------------|
| University of Central Florida                 | 54,513                             |
| Liberty University (VA)                       | 49,863                             |
| Texas A&M University—College Station          | 48,960                             |
| Ohio State University—Columbus                | 45,289                             |
| Florida International University              | 41,038                             |

First we will create a vector of enrollments.

```{r}
enrollments <- c()
```

Vectors can be combined. We will now create two different vectors. The first will contain enrollments from schools in Florida (named _florida_), while the other vector ( _other_ ) will contain enrollments from other schools.  Using c() with the vector names, we can concatenate both of these vectors into a new vector named _all.schools_

```{r}

```

We can do this faster by using *indexing*, using the syntax variable_name[index].  Note: the first element in a vector is element 1, the second is element 2, etc.

```{r}

```

We can also run functions on vectors:

```{r}

```


# Data frames

Data frames are similar to a spreadsheet, and contain multiple observations (rows) and variables (columns). 

For this class, all data are stored in .csv format. Let's load enrollments.csv. There are two ways we can do this.

A) Download the .zip file or sync git. Then, in your local directory, read in the file

```{r}
enrollments <- read.csv("enrollments.csv")
```

B) "Messy" Download the file from the internet into your local directory
```{r}

# download.file( ,"./enrollments.csv")
enrollments <- read.csv("enrollments.csv")
```

We can see the variable names in our data frame using _names()_

```{r}

```

We can select columns (vectors) in a dataframe by name using the *$* operator.  When using the *$* operator we place the name of the data frame to the left and the name of the variable to the right.

Let's pull out a vector for the variable _enrollment_

```{r}

```

To create a new column in our data frame, we can also use the *$* operator

```{r}

```


## Group work

Let's assume that class sizes  (freshmen, sophomores, etc) are equal. Create a variable containing the size of the freshman class at each school. 

```{r}

```

Since we are talking about people, we should not have decimals. Update the code above to use the floor() function to round down your answer to the nearest integer.

```{r}

```

What is the average size of the freshman class within the dataset?

```{r}

```


## If time remaining: practice with Swirl

Run the following. Note: you must be connected to the internet

```{r}
install.packages("swirl")
library(swirl)
install_course_github("kosukeimai","qss-swirl")
```

Then, in the console (bottom left window), type: swirl()
Follow the prompts, and select the course 'qss-swirl'. Then choose Selection 1 (Intro 1). Continue until you get to 80% completion.

Remember to type bye() in the console to leave Swirl. 

## Evaluation

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
