---
title: "Class 20"
author: " "
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
- Confidence Intervals Review
- Group work
- t-statistics (lecture)
- t-statistics in R

# Review: qnorm()

```{r}
# Normal distribution (don't need to know this code)
x <- seq(-4, 4, length=100)
hx <- dnorm(x)
plot(x,hx,lty=1,type="l")

# 90% confidence interval

#abline(v=qnorm(),col="burlywood")
#abline(v=qnorm(),col="burlywood")

# 95% confidence interval

#abline(v=qnorm(),col="thistle")
#abline(v=qnorm(),col="thistle")

# 99% confidence interval

#abline(v=qnorm(),col="darkgoldenrod")
#abline(v=qnorm(),col="darkgoldenrod")

```

# Review of group work

Reload the Boston immigration experiment dataset. The dependent variable is "diff_imm", which measures the change in support for immigration. The treatment variable is called _treatment_

```{r}
library(tidyverse)
boston <- read.csv("data/boston.csv")
boston$diff_imm <- boston$immigration_after - boston$immigration_before
```

2.1 Calculate the average treatment effect, as well as the standard error. Remember that since this is a difference in means between two samples (treatment and control), you will need to use this formula to calculate the standard error:

sqrt( (sd_treatmentˆ2 / n_treatment) + (sd_controlˆ2 / n_control)) 

```{r}
n_treatment <- 
n_control <-
sd_treatment <- 
sd_control <-

```

2.2 Generate a 99% confidence interval for the average treatment effect, and interpret the results.

```{r}

```

[Answer here]



# Standard error and regression

For a _binary treatment indicator_, standard errors can be quickly calculated using linear regression. Let's regress the treatment variable on the diff_imm variable.

```{r}

```

We can use _confint()_ to quickly calculate confidence intervals for regression coefficients.

```{r}

```


# Group Work: Confidence Intervals

Let's re-evaluate the study on racial bias in hiring. 

```{r}
resume <- read.csv("data/resume.csv")
```

In this experiment, identical resumes were sent out, except for the signaled race and gender, which was randomly assigned. We are interested in whether a resume received a call back from an employer or not.

Variables:
race    Signaled race on resume
sex     Signaled gender on resume
call    A binary variable indicating a call back (1) or not (0) 

1.1 First, calculate the average treatment effect:

Avg (Callback for White Resumes) - Avg (Callback for Black Resumes) 

```{r}
```

1.2 Now calculate the standard error for this difference in means. You can either use this formula:
sqrt( (sd_treatment^2 / n_treatment) + (sd_control^2 / n_control)) or a regression.


```{r}
n_treatment <- 
n_control <-
sd_treatment <- 
sd_control <-
```

1.3 Calculate a 90% confidence interval for the difference in means. 

```{r}

```

1.4 With 90% confidence, can we conclude based on 1.3 that there is racial bias in hiring?

1.5: Would this change with a 99% confidence interval (critical value = 2.58)?

```{r}

```

1.6 Conceptual question:
The 90% confidence interval for the average of Group A is -0.04 to 0.07
The 90% confidence interval for the average of Group B is -0.03 to 0.17

Can we conclude that B > A?

# Lecture: t-distributions and t-statistics

## Demonstrations in R

Compare t to normal (don't need to know this code)
```{r}
# Do not need to know this code
n <- samplesize <- 5
x <- seq(-4, 4, length=100)
hx <- dnorm(x)
plot(x,hx,lty=1,type="l",col="blue")
lines(x, dt(x,n-1), lwd=2, col="red")
```

Manual calculation of a t-statistic:

Mean:          .19
Sample SD:   .3294
Sample Size:    20

```{r}

# First calculate the standard error

# Then calculate the t-statistic

```

Manual calculation of a t-statistic for a difference in means:

```{r}
chechen <- read.csv("data/chechen.csv")

result <- chechen %>% group_by(fire) %>%
summarize(out = mean(postattack), n = n(), variance = var(postattack))

diff <- result$out[2] - result$out[1]
se <- sqrt(result$variance[2]/result$n[2] + result$variance[1]/result$n[1])

# Calculate t-statistic (ate/se)


```

What does a t-statistic of -1.33 mean?

```{r}
# Do not need to know this code
n  <- 318
x <- seq(-4, 4, length=100)
plot(x,dt(x,n-1),lty=1,type="l")
abline(v=-1.33)
```

Note: for smaller sample sizes, we can use _qt()_ to obtain critical values and confidence intervals. This is the equivalent of _qnorm_ for the t-distribution.

```{r}
# qt( level, df)
```


_If Time_

The _t.test_ command can also rapidly generate t-statistics:

## One sample t-test (for a single distribution):

```{r}
chechen <- read.csv("data/chechen.csv")
#t.test(, mu = 0)
```

## Two sample t-test  (t-test for difference in means):

Chechen ATE t.test:

```{r}
#t.test(, , mu = 0)
```

Returning to the resume experiment, let's use a t.test to examine the hypothesis that the true racial discrimination in our sample is 0

```{r}

```

We can also use t-tests to check for accurate randomization. Let's check that sex was independent randomized:

```{r}

```



## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?


