---
title: "Class 10"
author: "Your name"
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* Quick Quiz Review
* Longitudinal vs Wide Data
* Correlation (lecture)
* Descriptive relationships (practice)

# Quick Quiz Review

```{r}
library(tidyverse)

download.file("https://raw.githubusercontent.com/resonance1/other/main/terrorism.csv","terrorism.csv")
terrorism <- read.csv("terrorism.csv") 

```

4. Then for the treated and control groups, calculate the median of the *median.income* variable, as well as the 10th and 90th percentiles. Are the distributions similar? 

```{r}

terrorism %>% group_by(treatment) %>% summarize(med=median(median.income, na.rm=T),
                                                q10=quantile(median.income,.1, na.rm=T),
                                                q90=quantile(median.income,.9, na.rm=T))
```


# Longitudinal vs wide data

```{r}
austria <- read.csv("data/austria.csv") %>% unique()

```

Moving from long -> wide:

Function: pivot_wider(names_from = TIME, values_from = OUTCOME,names_prefix="your prefix")

```{r}


```

Moving from wide -> long:

Function: pivot_longer(cols=starts_with("your prefix"), names_to=TIME,values_to=OUTCOME,names_prefix="your prefix")

```{r}

```

Graphing long data

```{r}

treated <- austria %>% filter(treated==1) %>% group_by(year) %>% summarize(outcome=mean(socialist_voteshare))
control <- austria %>% filter(treated==0) %>% group_by(year) %>% summarize(outcome=mean(socialist_voteshare))


```


# Lecture: Correlation

# Exploring descriptive relationships in data

Descriptive correlations can provide helpful information to policymakers (remember that any causal relationship between two variables will have a non-zero correlation). In this exercise, you will evaluate the factors that are correlated with violent crime, using data from Denmark.


```{r}
denmark <- read.csv("data/denmark.csv")
```

Variables:

violentcrime_rate          Violent crime (scaled by population)
murderrate                 Murders (scaled by population)
lowincome_rate             Poverty Rate (scaled by population)
unemp_rate                 Unemployment Rate (scaled by population)
pct1565                    % of Population Aged 15-65 (scaled by population)
pctforeign                 % Non-citizen population (scaled by population)
sd_share                   Social Democratic Share of council seats
pop                        Population
eEducationhead             Spending on education, per capita
eHousinghead               Spending on public housing, per capita
eSocialhead                Spending on social expenditures, per capita
eTrafficInfrastructure     Spending on infrastructure, per capita
eTotalhead                 Total spending, per capita
drug_total                 Total drug offenses
murder                     Total murders
violentcrime               Total violent crimes


Let's explore potential correlations between violent crime and other municipal-level variables using a scatter plot.

```{r}
y <- denmark$violentcrime_rate
x <- 
  
plot()
```

We can add a trend line to the plot using lines(lowess(x,y))

```{r}
```

We can also evaluate the correlation numerically using cor(x,y,use="complete.obs"):

```{r}
```

With your neighbors, explore other correlations. Which factors are most strongly correlated with violent crime? Use a separate code chunks for each variable you explore. 

```{r}
```

```{r}
```

```{r}
```


If time remaining: work on PSET 4


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?

