---
title: "Homework 1"
author: ""
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

_Note_: You are welcome to work on the problem set before Friday's class. However, note that by the end of Friday's class we will cover all of the material you will need to complete the problem set.

# Bias in Self-reported Turnout

Surveys are frequently used to measure political behavior such as voter turnout, but some researchers are concerned about the accuracy of self-reports. In particular, they worry about possible social desirability bias. In this phenomenon, respondents who did not vote in an election tend to report that they did. Is such a bias present in the American National Election Study (ANES)? 

The ANES is a nation-wide survey that has been conducted for every election since 1948. The ANES conducts face-to-face interviews with a nationally representative sample of adults.  The following variables are in the turnout.csv file:

| Variable    | Description                                             |
|-------------|---------------------------------------------------------|
| year        | election year                                           |
| ANES        | ANES estimated turnout rate                             |
| VEP         | Voting Eligible Population (in thousands)               |
| VAP         | Voting Age Population (in thousands)                    |
| total       | total ballots cast for highest office (in thousands)    |
| overseas    | total eligible overseas voters (in thousands)           |


## Loading the data

You only need to use one of the following options.

Sync with git, or download the zip file and place "turnout.csv" into your main directory. In other words, the turnout.csv file should be visible within the directory view in the bottom right (the file browser). Then run the code below:

```{r}
turnout <- read.csv("turnout.csv")
```

## Questions

### Question 1

1.1. How many observations (rows) and variables (columns) are there in the data set? Use R functions. 

```{r}

```

1.2. What is the range of years covered in this data set? Note: you can look up basic R functions here: https://cran.r-project.org/doc/contrib/Short-refcard.pdf

```{r}

```

### Question 2

2.1. Using the *$* symbol to reference specific columns in the dataset, create a new variable that contains the turnout rate for each year as a proportion of the voting eligible population (VEP). Calculate the turnout rate for each year in the data set using the voting eligible population or VEP.  Double click on the following code to see the formula. $$VEP\: turnout\: rate = \frac{total\: votes}{voting\: eligable\: population}$$

```{r}

```

2.2. What is the average turnout rate in the data? What is the median turnout rate? Functions to use: *mean()*, *median()*

```{r}

```


### Question 3

Now we will examine social desirability bias.

3.1. Create a new variable named *vep_anes_diff* that computes the difference between the actual turnout rate (calculated in 2.1) and ANES estimates of the turnout rate for each year.

```{r}

```

3.2. How large is the difference on average? What is the range of the difference? 

```{r}

```

3.3. What do these results suggest?

[Answer in full sentences here]


Please Knit and submit as a .PDF using Canvas
