---
title: "Problem Set 3"
author: ""
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
library(tidyverse)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Are assassinations a natural experiment?

One longstanding debate in the study of international relations concerns the question of whether individual political leaders can make a difference. Some emphasize that leaders with different ideologies and personalities can significantly affect the course of a nation. Others argue that political leaders are severely constrained by historical and institutional forces.  Did individuals like Hitler, Mao, Roosevelt, and Churchill make a big difference?  The difficulty of empirically testing these arguments stems from the fact that the change of leadership is not random and there are many confounding factors to be adjusted for.

In this exercise, we consider a *natural experiment* in which the authors argue that the success or failure of assassination attempts is essentially random.

```{r}
library(tidyverse)
leaders <- read.csv("leaders.csv")

```


| Name           | Description                                                                      |
|----------------|----------------------------------------------------------------------------------|
| country        | The name of the country                                                          |
| year           | Year of assassination                                                            |
| leadername     | Name of leader who was targeted                                                  |
| age            | Age of the targeted leader                                                       |
| politybefore   | Average polity score during the 3 year period prior to the attempt               |
| polityafter    | Average polity score during the 3 year period after the attempt                  |
| civilwarbefore | 1 if country is in civil war during the 3 year period prior to the attempt, or 0 |
| civilwarafter  | 1 if country is in civil war during the 3 year period after the attempt, or 0    |
| interwarbefore | 1 if country is in international war during the 3 year period prior to the attem |
| interwarafter  | 1 if country is in international war during the 3 year period after the attempt, |
| result         | Result of the assassination attempt, one of 10 categories                        |
| result_binary  | 1 if the assassination resulted in the leader's death, 0 otherwise.              |

The polity variable represents the so-called *polity score* from the Polity Project.  The Polity Project systematically documents and quantifies the regime types of all countries in the world from 1800.  The polity score is a 21-point scale ranging from -10 (hereditary monarchy) to 10 (consolidated democracy).  

1.1 What was the average polity score of countries where assassination attempts occurred, before the attempt? Were any leaders of fully consolidated democracies (10 on the scale) the targets of an assassination attempt?

```{r}

```

1.2 What proportion of attacks resulted in the leaders' death? 

```{r}

```

1.3 Is a successful assassination attempt linked with further democratization or a descent into dictatorship?  To assess this, evaluate whether the *change* in polity scores is different for countries that faced successful vs. unsuccessful assassination attempts.

```{r}

```

[Full sentence(s) here]

1.4 Let's examine balance between the treated and control group in this natural experiment. What is the average polity score before an attempt, for successful and non-successful assassinations? What is the average age of the targeted leader? What is the proportion of countries that were engaged in a civil or international war in the 3 years prior to the attempt?

```{r}

```

1.5 Let's visualize some of these differences. Create boxplots of age, separately for each group

```{r}

```

1.6 Given the results above, evaluate the quality of this natural experiment. 

[Full sentences here]



# Poverty and Decision-Making

Do changes in one's financial circumstances affect one's decision-making process and cognitive capacity? In an experimental study, researchers randomly selected a group of US respondents to be surveyed before their payday and another group to be surveyed after their payday. Under this natural experiment design, the respondents of the `Before Payday` group are more likely to be financially strained than those of the `After Payday` group.  The researchers were interested in investigating whether or not changes in people's financial circumstances affect their decision making and cognitive performance.  

In this study, the researchers administered a number of decision-making and cognitive performance tasks to the `Before Payday` and `After Payday` groups. We focus on the *numerical stroop task*, which measures cognitive control.  In general, taking more time to complete this task indicates less cognitive control and reduced cognitive ability. They also measured the amount of cash the respondents have, the amount in their checking and saving accounts, and the amount of money spent.  The data set is in the CSV file `poverty.csv`. The names and descriptions of variables are given below:

-------------------------------------------------------------------------------
 Name                         Description
 ---------------------------- ---------------------------------------------------------
  `treatment`                 Treatment conditions: `Before Payday` and `After Payday`
  
  `cash`                      Amount of cash respondent has on hand
  
  `accts_amt`                 Amount in checking and saving accounts
  
  `stroop_time`               Log-transformed average response time for cognitive 
                              stroop test
  
  `income_less20k`            Binary variable: `1` if respondent earns less than 20k a 
                              year and `0` otherwise
-------------------------------------------------------------------------------
 
```{r}
poverty <- read.csv("poverty.csv")
```

## Question 1

2.1 Let's first rescale the cash and accts_amt variables so that the information is presented in thousands of dollars (i.e. divide each by 1000 and save as a new variable). Save these variables as cash_scaled and accts_scaled within the dataset.

```{r}

```

2.2 Now, let's examine the primary outcome of interest for this study -- the effect of a change in financial situation (in this case, getting paid on payday) on economic 
decision-making and cognitive performance. Begin by calculating the average treatment effect for the `stroop_time` variable (a log-transformed variable of the average response time for the stroop cognitive test). Interpret the resuts. Note that the difference between two logged values can be interpreted as a percent change.

```{r}

```

2.3 Let's take a closer look at the assumptions of this natural experiment -- namely whether or not the `Before Payday` versus `After Payday` treatment created measurable differences in financial circumstances. Compare the mean of cash_scaled and accts_scaled, separately for individuals in the `Before Payday` versus `After Payday` condition. Briefly interpret the results.

```{r}

```

2.4 Now create histograms for cash_scaled and accts_scaled (for all individuals). Are the distributions left-skewed, centered, or right-skewed?

```{r}

```

2.5 When variables are skewed, the median may be a more appropriate measure than the mean (which is susceptible to outliers). Recalculate your answer to 2.4 using the median difference between the two groups rather than the average difference. Thinking about the scale the variable is measured on, is the median difference between the `Before Payday` and the `After Payday` treatment large or small in substantive terms? What does this suggest about the author's research design?

```{r}


```

2.6 One possibility is that the assumptions of the research design may hold for lower-income earners, who are more likely to live paycheck to paycheck. Calculate the treatment effect on acct_scaled separately for respondents earning more than 20,000 dollars a year and for those earning less than 20,000 dollars a year, using both the average difference and the median difference. Briefly interpret the results.

```{r}


```

2.7 Lastly, recalculate 2.2 using only the group earning less than 20,000 dollars. Is the researcher's hypothesis that financial stress impacts cognitive capacity borne out for this low-earning group? 

```{r}

```


Please submit as a .PDF using Canvas